---
title: Going beyond mean effect size; presenting prediction intervals for on-farm
  network trial analyses - Supplementary Material
author: "Anabelle Laurent, Peter Kyveryga, Fernando Miguez, David Makowski"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
  html_document: default
  fig_caption: true
---

Required R packages and corresponding version 
```{r}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(dplyr)     # version 0.8.3
library(ggplot2)   # version 3.2.1
library(MCMCglmm)  # version 2.29
library(lme4)      # version 1.1-21
library(Hmisc)     # version 4.2-0
library(forcats)   # version 0.4.0
```

```{r}
## You might want to adjust the path as needed
simdata<-read.csv("../data/simdata.csv")
```

For the purpose of the supplementary material, we used a data frame called "simdata" including simulated yield ratio (lrr) of a hypothetical management practice to a control for each replicate (rep) within one trial identifier (Trial_ID).
Please find the csv file under the folder called data

```{r}
head(simdata)
```

# Frequentist approach using lme4 R package (Bates et al., 2015)

First, run the random-effect model 
```{r}
mod_fq <- lmer(lrr ~ 1  + (1|Trial_ID), data = simdata) 
```

* point estimate of log ratio (median)
```{r, message = FALSE}
fit <- fixef(mod_fq)                         
```

* Lower bound of confidence interval (2.5%)
```{r, message = FALSE}
low <- as.data.frame(confint(mod_fq))[3,1]
```

* Upper bound of confidence interval (97.5%)
```{r, message = FALSE}
up <- as.data.frame(confint(mod_fq))[3,2]
```


* prediction interval based on the method of Higgins et al. (2009) 

```{r}
bs_var <- (as.data.frame(VarCorr(mod_fq))[1,5])^2 # estimated between-trial variance 
se <- as.data.frame(coef(summary(mod_fq)))[1,2]  # standard error of the intercept 
var <- se^2                                      # variance of the estimated mean effect
k <- length(unique(simdata$Trial_ID))            # number of trials
mu <- fixef(mod_fq)                                # quantile 0.5 of the prediction interval
low <- mu - qt(.975, df=k-2) * sqrt(bs_var+var)      # quantile 0.025 of the prediction interval
up <- mu + qt(.975, df=k-2) * sqrt(bs_var+var)       # quantile 0.975 of the prediction interval
```


* probability of ineffective treatment 

```{r, echo=T, results='hide'}
var_total <- sqrt(bs_var+var)  # total variability 
q <- mu/var_total              # mean = q*total variability 
pt(q, df=k-2)                # student t distribution
proba <- 1 - pt(q, df=k-2)        # probability of ineffective treatment 
```


# Bayesian approach  using MCMCglmm R package (Hadfield, 2010)

* credibility interval and individual trial yield response  

First, run the random-effect model 
```{r}
prior1 <- list(B = list(mu = 0,V = 2),              # prior on the fixed effect
               G = list(G1 = list(V = 1, nu = 0.002)),  # prior on the between-trial variance
               R = list(V = 1, nu = 0.002))     # prior on the within-trial variance

mod_bayes <- MCMCglmm(lrr ~ 1,      # fixed effect 
                      random = ~Trial_ID,  # random effect
                      data = simdata,
                      family = "gaussian", 
                      prior = prior1,      # priors (defined above)
                      thin = 10,           # thinning interval
                      nitt = 200000,      # number of iterations
                      burnin = 20000,    # burnin
                      verbose = FALSE,     # if true MH diagnostics are printed to screen
                      pr = TRUE,           # posterior distribution of random effects are saved
                      DIC = TRUE)          # deviance information criterion
```


* credibility interval of the mean log ratio
```{r}
cred_int <- predict(mod_bayes, interval="confidence",level=0.95)[1,]
```

fit represents the point estimate of log ratio (median)  
lwr represents the lower bound of credibility interval (quantile 0.025)    
upr represents the upper bound of credibility interval (quantile 0.975)  

* credibility interval of individual log yield ratio  

```{r}
pred <- predict(mod_bayes, interval="confidence", marginal = mod_bayes$Trial_ID, level = 0.95)
pred <- as.data.frame(pred)
pred$trial <- simdata$Trial_ID 

fit <- pred$fit
lwr <- pred$lwr
upr <- pred$upr
pred$trial <- reorder(pred$trial, pred$fit)

duplication <- which(duplicated(pred))
pred <- pred[-duplication,]
print(pred)    # dataframe returning the credibility interval of individual log ratio
```

The column "fit" represents the point estimate of log ratio for each individual trial (see column "trial") (median)  
The column "lwr" represents the lower bound of credibility interval for each individual trial (quantile 0.025)  
The column "upr" represents the upper bound of credibility interval for each individual trial (quantile 0.975)  

* prediction interval for a new trial  

In order to compute a prediction interval, you need to add a new trial with missing value was added (here called IDNew) and run the random effect model.  

```{r}
simdata <- simdata %>% add_row(Trial_ID = "IDNew")
tail(simdata)
```

```{r}
mod_bayes_new <- MCMCglmm(lrr ~ 1,  # fixed effect 
                          random = ~Trial_ID,  # random effect
                          data = simdata,
                          family = "gaussian", 
                          prior = prior1,      # priors (defined above)
                          thin = 10,           # thinning interval
                          nitt = 200000,      # number of iterations
                          burnin = 20000,    # burnin
                          verbose = FALSE,     # if true MH diagnostics are printed to screen
                          pr = TRUE,           # posterior distribution of random effects are saved
                          DIC = TRUE)
```

Instead of using the function predict, the MCMC chains were used to compute the prediction interval 
```{r}
REnew <- as.data.frame(mod_bayes_new$Sol[,"Trial_ID.IDNew"]) # MCMC chain for the random effect IDNew
intercept <- as.data.frame(mod_bayes_new$Sol[,"(Intercept)"]) # MCMC chain for the the intercept
newtrial <- REnew + intercept                                # sum the two MCMC chains 
predinterval <- quantile(newtrial[,1], c(.025, .5, .975))  # get quantiles 
```

* probability of ineffective treatment

```{r}
X <- sum(newtrial$var1 < 0)   # number of cases where log ratio <0
n <- length(newtrial$var1)  # total number of possible cases 
proba <- X/n                # probability of ineffective treatment 
```

You do need an exponential transformation to convert the Frequentist and Bayesian outputs from log yield ratio to yield ratio. 

## Figures
We use the exponential transformation for the figures 
```{r}
# outputs = dataframe combining trial estimations, credible interval and prediction interval
outputs <- pred %>% 
  add_row(fit = cred_int[1],
          lwr = cred_int[2],
          upr = cred_int[3],
          trial = "mean yield ratio") %>%
  add_row(fit = predinterval[2],
          lwr = predinterval[1],
          upr = predinterval[3],
          trial = "prediction interval")
  
outputs <- outputs %>%
  mutate(trial = forcats::fct_reorder(trial,fit)) %>%
  mutate(trial = forcats::fct_relevel(trial, "prediction interval", after = 0)) %>%
  mutate(trial = forcats::fct_relevel(trial, "mean yield ratio", after = 1)) 
```


```{r, fig.width=5, fig.height=7, fig.cap = "Estimated individual effects (denoted by T and a numerical value) with their 95% credible intervals, mean yield ratio and its 95% credible interval, and 95% prediction interval. The numerical value at the bottom-right indicates the probability of ineffective treatment (probability of yield ratio less than 1) in a new trial"}
ggplot(outputs, aes(x = exp(fit), xmin = exp(lwr), xmax = exp(upr), y = trial )) +
  geom_point(size = 2) +
  geom_segment( aes(x = exp(lwr), xend = exp(upr), y = trial, yend=trial)) +
  theme_bw() + 
  xlab("Yield ratio") + 
  ylab("") +
  geom_text(size = 5,aes(x = 5.2,y = 1.2, label = ifelse(trial=="prediction interval", 
                                                         round(proba,2),""))) +
  geom_vline(xintercept = 1, color = "black", size = 0.8) +
  theme(axis.text = element_text(size = 12, face = "bold"), 
        axis.title = element_text(size = 12, face = "bold"), 
        legend.text = element_text(size = 12, face = "bold"),
        legend.title = element_blank())
```


### References

Bates, D., Mächler, M., Bolker, B., Walker, S., 2015. Fitting Linear Mixed-Effects Models Using lme4. Journal of Statistical Software 67. https://doi.org/10.18637/jss.v067.i01  


Hadfield, J.D., 2010. MCMC Methods for Multi-Response Generalized Linear Mixed Models: The MCMCglmm R Package. Journal of Statistical Software 33. https://doi.org/10.18637/jss.v033.i02

Higgins, J.P.T., Thompson, S.G., Spiegelhalter, D.J., 2009. A re-evaluation of random-effects meta-analysis. Journal of the Royal Statistical Society: Series A (Statistics in Society) 172, 137–159. https://doi.org/10.1111/j.1467-985X.2008.00552.x